---
title: "Lab 07 - Conveying the right message through visualisation"
author: "Jamieson Nathan"
date: "04/09/2025"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

### Exercise 1

```{r FirstRun}

covid_data <- tribble(
  ~date,       ~cases_per_100k, ~mask_mandate,
  "2020-07-01", 23,             "Mask",
  "2020-07-01", 14,             "No Mask",
  "2020-07-15", 20,             "Mask",
  "2020-07-15", 12,             "No Mask",
  "2020-08-01", 19,             "Mask",
  "2020-08-01", 10,             "No Mask",
  "2020-08-15", 17,             "Mask",
  "2020-08-15", 8,              "No Mask",
  "2020-08-20", 16,             "Mask",
  "2020-08-20", 7,              "No Mask"
)

print(covid_data)

```

### Exercise 2

``` {r SecondRun}

library(ggplot2)
library(tibble)

covid_data$date <- as.Date(covid_data$date, "%Y-%m-%d")

ggplot(data = covid_data, aes(x = date, y = cases_per_100k, color = mask_mandate, group = mask_mandate)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Mask" = "orange", "No Mask" = "blue")) +
  labs(title = "7-Day Rolling Average of Daily COVID-19 Cases per 100k Population",
       subtitle = "Comparing Mask vs. No-Mask Mandate Counties in Kansas",
       x = "Date",
       y = "Cases per 100k Population",
       color = "Mandate Status") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5))


```

Based on this visual, which ahs approximately the same shape of data, it is clear that the no mask counties have has lower rates of COVID-19 cases. 

The discrepency in visuals show how rife misinformaion and politicizing data was during this period! Although we have not tested whether the difference is significant, this shows at a face-value that no mask counties had lower rates. I could see this being explained in a few ways: 1) no mask counties had another variable that made them less suceptible (lower population density, more rural, etc.), 2) masks somehow increase rates of COVID (less likely, if anything I would think rates would be comparable), or 3) Rates of testings were higher in mask-mandate counties. 

``` {r FourthRun}

ggplot() +
  geom_line(
    data = subset(covid_data, mask_mandate == "Mask"),
    aes(x = date, y = cases_per_100k, color = "Mask"),
    size = 1.2
  ) +
  geom_line(
    data = subset(covid_data, mask_mandate == "No Mask"),
    aes(x = date, y = cases_per_100k * 3, color = "No Mask"),
    size = 1.2
  ) +
  scale_color_manual(
    name = "Mandate Status",
    values = c("Mask" = "orange", "No Mask" = "blue")
  ) +
  scale_y_continuous(
    name = "Mask",
    limits = c(0, max(covid_data$cases_per_100k) * 2),
    sec.axis = sec_axis(~ . / 3, name = "No Mask")
  ) +
  labs(
    title = "Misleading Plot",
    subtitle = "Likely a CNN Graphic",
    x = "Date"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

```

The key difference between these two is the differing Y-Axes, such that at face-value the no-mask line appears to be higher, even though they are not corresponding to the same scale. 

This new visual conveys an even more dramatic difference in lines, and to the untrained eye it appears as it 'should', with the no mask counties having higher rates. 
